### 2012-06-08 RateEstimation.tex
.tex of Jon's Rate Estimation document

%\documentclass[12pt]{iopart}
\documentclass[prd,aps,eqsecnum]{revtex4}
\usepackage{epsfig,graphicx,latexsym}
\def\rmd{{\rm d}}
\def\msun{$M_{\odot}$}
\def\MNRAS{{\em Mon. Not. Roy. Astron. Soc. }}
\def\CQG{{\em Class. Quantum Grav. }}
\def\PRD{{\em Phys. Rev. D }}
\def\apj{{\em Astrophys. J.}}
\newcommand{\erf}[1]{(\ref{#1})}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\rmd{{\rm d}}

\begin{document}

\title{Rate Estimation}
%\author{Jonathan R Gair}
%\address{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK}
%\email{jgair@ast.cam.ac.uk}
\author{Will, Jon and Ilya}

\date{\today}

\begin{abstract}
Some notes on the discussion Will and I had today (March 23rd 2012).
\end{abstract}

\maketitle

\section{Rate Estimation}
\subsection{Likelihood}
The data we observe is a sequence of $N$ events with statistic values $x_i$, above some threshold $x_0$. These events could be drawn from a background described by the rate distribution $\nu_b(x)$, which specifies the rate of events arriving with statistic value greater than $x$, or from a foreground described by the rate distribution $\nu_f(x)$. We use flags $f_i \in \{0,1\}$ to denote if an event if is foreground ($f_i=1$) or background ($f_i=0$). We want to use the data to estimate the $f_i$'s and some property of the rate distributions (we'll come back to this later).

The probability of seeing data $x = \{x_i\}$ {\it and} flags $f_i$ given the rates is the probability that we would draw events $\{x_j : f_j=1\}$ from the foreground distribution and the events $\{x_j : f_j=0\}$ from the background distribution. The probability of getting a set of $M$ events $\{y_i\}$ above threshold $y_0$ and with $y_1 < y_2 < \cdots y_M$ from a Poisson process with loudness distribution $\nu(y)$ is the probability of having $0$ events between $y_0$ and $y_1$, which is ${\rm e}^{-(\nu(y_0)-\nu(y_1))}$, times the probability of having an event at $y_1$, which is $-\nu'(y_1)$, times the probability of having no events between $y_1$ and $y_2$, which is ${\rm e}^{-(\nu(y_0)-\nu(y_1))}$, etc. Overall we find
\begin{equation}
p(\{y_i\}) = {\rm e}^{-\nu(y_0)} \prod_{i=1}^M [-\nu'(y_i)].
\end{equation}
Combining the foreground and background distributions we then have
\begin{equation}
p(x, \{f_i\} | \nu_f, \nu_b) = {\rm e}^{-\nu_f(x_0)} {\rm e}^{-\nu_b(x_0)} (-1)^N \prod_{\{j :   f_j=1\}} \nu_f'(x_j)\prod_{\{k : f_k=0\}} \nu_b'(x_k) .
\end{equation}
We want the posterior distribution for the parameters of the rate distributions and the $f_i$'s, which is
\be
p(\nu_f, \nu_b, \{f_i\} | x) \propto p(x, \{f_i\} | \nu_f, \nu_b) p(\nu_f, \nu_b)
\ee
The last term is just the prior probability of the rate parameters.

To relate this to Ilya's parameterisation, we say that the loudness distribution is determined by an overall scale, $\lambda$, and a probability distribution of events $p(x)$ with loudness greater than $x_0$, which we temporarily assume is known. Then we have 
\be
\frac{\rmd \nu}{\rmd x} = - \lambda p(x)
\ee
and $\nu(x_0) = \lambda$. The posterior probability then becomes
\be
p(\lambda_f, \lambda_b, \{f_i\} | x) \propto  p(\lambda_f, \lambda_b) \, {\rm e}^{-\lambda_f} {\rm e}^{-\lambda_b} \lambda_f^{N_f} \lambda_b^{N_b} \prod_{\{j :   f_j=1\}} p_f(x_j)\prod_{\{k : f_k=0\}} p_b(x_k)
\ee
where $N_f/N_b$ are the number of foreground/background events and $p_f/p_b$ are the corresponding loudness probability distributions. This is similar to what Ilya was writing down, except possible for the $\lambda_b$ and $\lambda_f$ dependent terms.

We can marginalise this probability distribution over the $f_i$'s which gives
\be
p(\lambda_f, \lambda_b | x) \propto p(\lambda_f, \lambda_b) \, {\rm e}^{-\lambda_f} {\rm e}^{-\lambda_b}  \prod_{j=1}^N (\lambda_f p_f(x_j) + \lambda_b p_b(x_j))
\ee
or we could marginalise over the rate parameters. This requires a specification of the prior on the rates. As an illustrative example, we can take a flat prior on the rate parameters and then we have
\be
p(\{f_i\} | x) \propto  N_f! N_b! \prod_{\{j :   f_j=1\}} p_f(x_j)\prod_{\{k : f_k=0\}} p_b(x_k).
\ee 
We can go further and work out the marginal distribution of a single $f_i$, but these distributions are not particularly pleasant. We can also look at 2D probabilities of pairs of $f_i$'s and would expect the see correlations. Thus, everything can be written down exactly, but given the number of possible combinations of $f_i$'s that we would have to evaluate, an MCMC will allow us to find the combined distribution more easily.
% e.g., for $f_1$ we have
%\be
%p(f_1=1|x) 
%\ee

\subsection{A few more thoughts}
The framework above allows us to estimate the form of the rate functions as well, i.e., the $p_f(x)$ and $p_b(x)$, if we assume some model for them. If we parameterize $p_f(x)$ by parameters $\{\mu^i_f\}$ and parameterize $p_b(x)$ by parameters $\{\mu^i_b\}$ then the combined posterior becomes
\be
p(\lambda_f, \lambda_b, \{\mu_f^i\}, \{\mu_b^i\}, \{f_i\} | x) \propto  p(\lambda_f, \lambda_b, \{\mu_f^i\}, \{\mu_b^i\}) \, {\rm e}^{-\lambda_f} {\rm e}^{-\lambda_b} \lambda_f^{N_f} \lambda_b^{N_b} \prod_{\{j :   f_j=1\}} p_f(x_j | \mu_f^i)\prod_{\{k : f_k=0\}} p_b(x_k | \mu_b^i) \label{fullpost}
\ee
so we can jointly recover the shape of the loudness distribution and the rate, if we assume some model. I think using the power-law or exponential models would be a good illustration even if it was not realistic for LIGO. However, we could also mimimc what Drew et al. do. If we divide the statistic space up into bins, we can then model $p_f(x)$ to be flat in each of these bins. The number of free parameters, $\{\mu_f^i\}$, is then the number of bins used. This is essentially what they do when they bin the data and look at the distribution of events from injections and background (timeslides).

Going one step further, we can use the injection and timeslide data directly in this framework. We just assume that we have another set of $N_I$ events, $\{ w_i \}$, known to be drawn from the foreground distribution and a set of $N_T$ events $\{z_i\}$ known to be dtawn from the background distribution. These events were not drawn with the correct rate parameters, but a fixed number of events were computed and so they do not contribute to the $\lambda_f$ and $\lambda_b$ terms, but they do contribute an extra factor
\be
\prod_{l=1}^{N_I} p_f(w_l | | \mu_f^i) \prod_{m=1}^{N_T} p_b(z_i | | \mu_b^i) 
\ee
on the right hand side of Eq.~(\ref{fullpost}). So, we could simultaneously fit for the shape of the background as well as the rates and the identification of individual sources. If we were using the binned model for the probability distributions we would presumably find that, if there were many more events in the timeslide and injection data set, this would reduce to the analysis that was described above which assumes the shape of the rate distributions are given by the distribution of those events. However, this framework would be more flexible in the limit where the amount of timeslide or injection data was limited.

Another thought was that the injection distribution might not reflect the astrophysical distribution of signals, in which case we might want to see what happens if we only use timeslide data and ignore injection data. Similarly, in the detection era we might not be able to estimate background very easily since there will be too any sources in the data set (we wish!). In that case maybe we would assume that foreground was reliable and background wasn't. All of this is easy enough within this framework, but we might want to think a bit more.

\end{document}